---
title: "FiveThirtyEight Net Favorability vs Election Margin Reproduced"
author: '@sakalli'
date: "1.10.2016"
output:
  html_document:
    theme: journal
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducing the analysis

Read this excellent analysis by Harry Enten at FiveThirtyEight <http://fivethirtyeight.com/features/trump-is-more-unpopular-than-clinton-is-and-that-matters/> and got curious about the results so did a quick reproduction. Was mostly interested in the R squared and p values for the regression, since with only 9 samples the material is extremely small.

## Conclusion

So as you can see below the adjusted R-squared is not bad at 0.79 and p-value at 0.0008 is good as well. Somewhat surprisingly good, as we are operating with only 9 observations. 

If we do assume that this model holds, it would predict that Clinton will win with a margin of a little more than 4%. Although the 95% prediction interval for the prediction is rather wide, from -6.2 to 14.5, which also means that a trump win is within it. If you lower the level to 62% you get the interval entirely on Clintons side. That means that as Clinton got the 9.8% advantage in favorability, the model **predicts that Clinton will win 62 out of a 100 times.** Which is not that different from what the FiveThirthyEight Polls Plus model forcasts at the moment.

If you have corrections or improvements please do a pull request at <https://github.com/skallinen/public>. 


## Calculations

```{r results='hide', message=FALSE, warning=FALSE}
# loading the libraries -
library(ggplot2)
library(dplyr)
library(rvest)
library(plotly)
```


```{r message=FALSE, warning=FALSE}

# loading the fivethirthyeigh data - 
url <- "http://fivethirtyeight.com/features/trump-is-more-unpopular-than-clinton-is-and-that-matters/"

favorability_table <- url %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="post-128864"]/div/section/table') %>%
  html_table()

favorability_table <-  favorability_table %>% data.frame()

names(favorability_table) <- favorability_table[1,]
favorability_table <- favorability_table[-1,]

# loading the election result data -
url2 <- "http://uselectionatlas.org/RESULTS/compare.php?type=national&year=2012&f=0&off=0&elect=0"

election_results <- url2 %>%
  read_html() %>%
  html_table(fill=T)

election_results <-  election_results[[1]] %>% data.frame()
names(election_results) <- election_results[2,]
election_results <- election_results[-1:-3,1:13] %>% select(Year, `%Margin`)


# combining the data -
combined_table <-  left_join(favorability_table, election_results, by = c("YEAR" = "Year"))

combined_table$DIFFERENCE <- gsub("[^\\d]+\\.?[^\\d]+","\\1",
                                  combined_table$DIFFERENCE, 
                                  perl = T) %>% as.numeric

combined_table$`%Margin` <- gsub("%","",combined_table$`%Margin`, 
                                 perl = T) %>% as.numeric

combined_table <- combined_table %>% 
  mutate(DIFFERENCE = ifelse(`WINNER PARTY`=="R",
                             DIFFERENCE * -1,
                             DIFFERENCE), 
                 `%Margin` = ifelse(`WINNER PARTY`=="R",
                                    `%Margin` * -1,
                                    `%Margin`)) %>% 
  filter(YEAR != 2016) %>% 
  rename(election_win_margin=`%Margin`, dem_favorability_advantage = DIFFERENCE)


# plotting the regression -
pl <- ggplot(combined_table, aes(dem_favorability_advantage, election_win_margin, label=WINNER)) + 
  geom_point() + 
  stat_smooth(method = "lm", colour = "#F07E26") + 
  xlab("Dem advantage in net favorability in late sept") + 
  ylab("Dem margin in election")

# customizing the colours -
pl <- pl + theme(
  panel.background = element_rect(fill = "#ECEAE6",
                                colour = "#DFDAD5",
                                size = 0.5, 
                                linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, 
                                  linetype = 'solid',
                                  colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, 
                                  linetype = 'solid',
                                  colour = "white")
  )

ggplotly(pl)        
# building the regression model - 
fit <- lm(election_win_margin ~ dem_favorability_advantage, data = combined_table)
summary(fit)

# predicting with the model -

# prediction interval level at 95%
predict(fit,  data.frame(dem_favorability_advantage=9.8),interval = "prediction", level = 0.95)

# prediction interval level at 62%
predict(fit,  data.frame(dem_favorability_advantage=9.8),interval = "prediction", level = 0.62)

```




<script type="text/javascript">(function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){ (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o), m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m) })(window,document,'script','https://www.google-analytics.com/analytics.js','ga'); ga('create', 'UA-74818-38', 'auto'); ga('send', 'pageview');</script>